\documentclass{psu-report}
\usepackage{psu-thesis-chicago}
\usepackage{todonotes}
\usepackage{amsmath,amssymb,amsfonts,amsthm,mathtools,mathrsfs}
\usepackage{listings}
\usepackage{minted}
\usepackage{siunitx}
\usepackage{tikz,pgfplots,pgfplotstable}
\usetikzlibrary{positioning, shapes.geometric}
\pgfplotsset{compat=newest}

\addbibresource{refs.bib}

\title{Organ Pipe Voicing}
\author{
    Meghana Cyanam \\
    Hai Nguyen \\
    Gabriel Pinochet-Soto \\
}
\doctype{report}
\psumonth{April}
\psuyear{2025}

\professorOne{Daniel Taylor-Rodriguez}
\professorTwo{Jacob Schultz}
\clientOne{Arno Patin}

\documentformat{monograph}

\newenvironment{code}{\begin{listing}}{\end{listing}}
\newcommand{\chapterauthor}[1]{%
  \vspace{-2em}
  {\large \textsc{#1}\par}
  \vspace{1.5em}
}

\begin{document}

\maketitle
\makecopyright

\begin{psuabstract}
    Intonation, or voicing, in pipe organ technology involves adjusting each
    pipe to produce its optimal sound.
    This process entails determining the appropriate tone that harmonizes with
    the organ’s style, the architectural setting, and the preferences of its
    owner.
    While the “perfect” tone is largely subjective, certain aspects, such as 
    volume and harmony, can be quantified using contemporary tools.
    
    The relationship between the diameter of the pipe’s toe, the size of the toe
    hole, airflow, and frequency is paramount in comprehending and optimizing
    organ pipe performance.
    These parameters interact to influence the pipe’s tonal characteristics,
    acoustic intensity, and harmonic structure. Adjustments in the toe hole size
    regulate airflow, affecting both frequency and tonal balance, while the pipe
    toe diameter significantly shapes the overall sound profile.
    Exploring these connections elucidates the intricacies of voicing and 
    scaling processes, providing insights into both traditional methods and
    contemporary approaches employing acoustic measurements.

    Our goal is to clarify the relationship between the Ising number and the
    parameters of the organ pipe, and to provide a model that can be used to
    dimension the pipes before the voicing stage.
\end{psuabstract}

% \begin{psudedication}
% \end{psudedication}

% \begin{psuacknowledgments}
% \end{psuacknowledgments}

\tableofcontents
\clearpage

\listoftables
\clearpage

\listoffigures
\clearpage

% \begin{psuglossary}
% \begin{description}
%     \item[LaTeX] A document preparation system
% \end{description}
% \end{psuglossary}

% \begin{psupreface}
% \end{psupreface}

\startbody

\chapter{Introduction}
\chapterauthor{Meghana Cyanam, Hai Nguyen, Gabriel Pinochet-Soto}

Pipe organs have been around for hundreds of years, and making them sound just
right is a special job called intonation, or \emph{voicing}.
This means adjusting each pipe to create the best sound for the organ.
The way a pipe looks and is built -- like its size, shape, and material-- can
change how it sounds.
These details are decided first, and then the pipes are fine-tuned before being
added to the organ.

Experts have been studying these details for a long time, and they’ve written
about different ways to make pipes sound beautiful and harmonious.
Our goal is to understand how the Ising number -- a special number that helps
with harmonization -- relates to the pipe’s design, how it behaves upon modification of
the airflow, and how we can use this information to:
\begin{enumerate}
    \item Make the Ising number more accurate.
    \item include the flow rate in the Ising number.
    \item Create a model that helps us design pipes before we start intonating them.
\end{enumerate}

\chapter{Action Plan}
\chapterauthor{Meghana Cyanam, Hai Nguyen, Gabriel Pinochet-Soto}

In this Chapter, we discuss different approaches for successfully improving the
organ pipe voicing process.
We discuss the project goals and objectives, potential solutions, and the
expected outcomes of the project.
Subsequently, we present the data available, some challenges and concerns that
we encountered.
Proposed methodologies are discussed and briefly explained.
Finally, we present the project timeline and the expected deliverables.

\section{Project Goals and Objectives}

The main goal of this project concerns the improvement of the organ pipe voicing
process.
As the organ pipe design is a complex process, and usually is not a reversible
process, it is important to ensure that the design of the pipes is done
correctly; this presents a key constraint for the intonation process.
The parameters involved are various, which makes learning about this complex
system a difficult task; the understanding of these parameters and their
interactions would facilitate the harmonization process.

A particular subgoal of this project is to improve the Ising
formula~\autocite{1971Ising-1}, which reads
\begin{equation}
    \label{eq:ising}
    \mathsf{I}
    =
    \frac{v}{\omega}\sqrt{\frac{d}{h^3}}
    =
    \frac{1}{\omega}\sqrt{\frac{2 P d}{\rho h^3}},
\end{equation}
where \(\mathsf{I}\) is the Ising number, \(\omega\) is the (target) frequency
of the pipe, \(v\) is the jet initial velocity, \(d\) is the jet initial
thickness, \(h\) is the cut-up height (or the length of the mouth), \(P\) is the
blowing pressure (where we require Bernoulli's Law, ~cf. \autocite{2025Lilj-1}),
and \(\rho\) is the density of the air.
From \autocite{1971Ising-1, 2025Lilj-1}, it is suggested that the Ising number
should be constant, preferably close to \( \mathsf{I} \approx 2\).

We would like to include the air flow rate in the Ising formula~\ref{eq:ising},
so this methodology can be used to dimension the pipes reliably,
before the voicing stage: as the amount of air flowing through the pipe has been
seen to affect the sound of the pipe, we would like to include this parameter in
the Ising formula~\ref{eq:ising}.

\subsection{Key questions}

The following questions are key to the project, and we aim to address them to
our best ability:
\begin{enumerate}
    \item Can the Ising formula~\ref{eq:ising} be improved, refined, or
        corrected, in order to match measured data?
    \item Can the flow rate be included in the Ising formula~\ref{eq:ising}?
    \item Can we obtain a model reliable enough, that can be used to dimension the
        pipes before the voicing stage?
\end{enumerate}

\section{Available data sources}

The main source of data is provided by the client.
The data is currently available in a \texttt{csv} file, and it contains the
following columns:
\begin{itemize}
    \item \textbf{isBourdon} \textit{boolean} -- Indicates if the pipe is a Bourdon pipe.
    \item \textbf{flueDepth} \textit{float} -- The depth of the flue.
    \item \textbf{frequency} \textit{float} -- The frequency of the pipe.
    \item \textbf{cutUpHeight} \textit{float} -- The cut-up height of the pipe.
    \item \textbf{diameterToe} \textit{float} -- The diameter of the toe.
    \item \textbf{acousticIntensity} \textit{float} -- The acoustic intensity of the pipe.
    \item \textbf{partialN} \textit{float} -- \(N\)th partial of the pipe.
        The number of partials is not fixed, and it can vary from 1 to 8.
        This value is bounded between 0 and 100.
\end{itemize}

\section{Methodology}

We intend to employ an integrated approach combining semianalytical methods, numerical methods, machine learning, 
and advanced statistical techniques to systematically address the client's question regarding optimal organ 
pipe intonation.
Each method is selected not only for its individual strengths, but also for how collectively they can provide 
a comprehensive understanding and practical solutions to the organ voicing process.

\subsection{Semianalytical methods}

As described in \autocite{2004RosFle-1}, the physics of the organ pipes relies
on the wave equation.
Exact solutions for the time-harmonic, longitudinal, wave equation can be
derived using the method of separation of variables.

These solutions give the eigenvalues (i.e., the frequencies). 
The study of these frequencies -- in particular, the fundamental
frequency -- is important for the intonation process.

We intend to explore basic semianalytical methods, such as the method of
separation of variables, to study the eigenvalues of the wave equation.
We rely on available literature~\autocite{2004RosFle-1, 2012RosFle-1}.

Specifically, our objectives include:
\begin{itemize}
    \item Review known formulas and estimations methods that connect the
        dimensions of the pipe with the harmonic frequencies of the pipe.
    \item Exhaust literature approaches to the Ising formula.
    \item Confirm that optimal voicing can be achieved by adjusting the
        dimensions of the pipe.
\end{itemize}

\subsection{Numerical methods: Finite Element Method}

The Finite Element Method (FEM) is a numerical method for solving partial
differential equations (PDEs) and integral equations.
As the governing equations of the organ pipe are PDEs, we can use FEM to solve
them.
Standard techniques are available, and we will not delve into the details of the
FEM.
See~\autocite{2021ErnGue-1, 2021ErnGue-2, 2019VazKeeDem-1} for a
comprehensive introduction to the FEM and PML methods.

Provided the geometry of the pipe, we can use the FEM to compute the frequencies
of the pipe, and intensity levels of the sound
(cf.~\autocite[Figure 1]{2009RucAugFia-1}).

The goals of FEM application are to:
\begin{itemize}
    \item Provide FEM simulations of the organ pipe, with the underlying
        physical model (time-harmonic wave equation).
    \item Compare FEM simulations with the Ising formula and the client-provided data.
    \item Allow artificial data generation.
    \item Compute the intensity (i.e., the sound level) of the simulated pipe,
        and determine if the harmonics are ``adequate'' (i.e., if they are
        ``good sounding'').
\end{itemize}

\subsection{Machine learning techniques}

Machine learning techniques, specifically neural networks, will leverage numerical simulations and empirical 
client-provided data to predict optimal pipe dimensions and acoustic properties. 
A neural network with tailored architecture—including a Softplus activation function for predicting the 
Ising number and sigmoid functions for partial ratios—will be trained. 
This method will effectively capture complex nonlinearities and relationships inherent in organ pipe acoustics.

% Our goals are to:
The machine learning objectives are:
\begin{itemize}
    \item provide a neural network model that can predict the Ising number and
        the partials ratios.
    \item compare the neural network predictions with the Ising formula and the
        data provided by the client.
\end{itemize}

\subsection{Generalized Additive Models}

Generalized Additive Models (GAMs) offer a robust statistical framework for modeling complex, 
nonlinear interactions between multiple pipe parameters such as toe-hole diameter (flow rate), flue dimension,
cut-up height, and blowing pressure—parameters critically affecting pipe tonal quality. 
GAMs will integrate empirical data, facilitating a nuanced understanding of how these factors interact in a nonlinear 
manner, thus providing a data-driven approach to refine existing formulas, like the Ising formula, and support 
practical decision-making in pipe manufacturing and voicing.

Our GAM-specific goals include:
\begin{itemize}
    \item Constructing flexible predictive models that accurately determine the Ising number and partials ratios.
 \item Comparing GAM predictions to both analytical models and real-world data.
 \item Providing actionable insights for pipe dimensioning ahead of the voicing stage, enhancing efficiency 
    and tonal consistency in organ manufacturing.
\end{itemize}

\section{Project timeline}

The project timeline is the following:
\begin{itemize}
    \item Weeks 1-2: Data acquisition, cleaning, and preliminary analysis.
    \item Weeks 3-4: Preliminary exploration of the data.
    \item Weeks 5-8: Implementation of the methodologies.
    \item Weeks 9: Synthesis of the results.
    \item Weeks 10: Final report and presentation.
\end{itemize}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Current work!

\chapter{Exploratory Data Analysis}

\chapterauthor{Hai Nguyen, Gabriel Pinochet-Soto}
% From Hai's work

We first start with the process of data wrangling.
We do not include constants or parameters that are given by the client --say,
the blowing pressure, the density of the air, or the jet initial velocity-- in
the Ising number formula~\ref{eq:ising}.

Rather, we focus on the observable parameters that can be measured, as described
above.
Recall that these quantities are:
\begin{itemize}
    \item \textbf{flueDepth} -- The depth of the flue.
    \item \textbf{frequency} -- The frequency of the pipe.
    \item \textbf{cutUpHeight} -- The cut-up height of the pipe.
    \item \textbf{diameterToe} -- The diameter of the toe.
    \item \textbf{acousticIntensity} -- The acoustic intensity of the pipe.
    \item \textbf{partialN} -- The \(N\)th partial of the pipe.
\end{itemize}
We perform standard computations to retrive the Ising number and the flow rate
from the data.

\section{Gradient Boosted Decision Trees}

We use Gradient Boosted Decision Trees (GBDT) to study the relationship
between the Ising number and the parameters of the organ pipe.
GBDT~\autocite{2001Fri-1} is implemented in Scikit-learn~\autocite{2016Kra-1}.
This method is used for predicting continuous values (like frequency
distributions).
It learns patterns in the data by combining multiple decision trees in a
sequential way, called boosting.
Once trained, the model evaluates which features (geometric parameters)
contribute the most to predicting the frequency distribution.
We consider the \textbf{flueDepth}, \textbf{cutUpHeight}, and
\textbf{diameterToe} as the independent variables, and the
partial distribution as the dependent variable.
We can confirm that cut-up height is an relevant factor in the Ising formula.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            xlabel={Flow rate},
            ylabel={Predicted Ising number},
            grid=major%
        ]
        \addplot table[x=flowRate, y=predictedIsing, col sep=comma, only marks] {../../Data/flowIsingData.csv};
        \end{axis}
    \end{tikzpicture}
    \caption{Flow rate vs. predicted Ising number.}
\end{figure}

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            xlabel={Frequency (Hz)},
            ylabel={Cut-Up Height (m)},
            colorbar,
            colormap/hot,
            point meta=explicit,
            grid=major
        ]
        \pgfmathsetmacro\jitter{rand*0.0005}
        \addplot[
            scatter,
            only marks,
            mark size=1.5pt,
            opacity=0.7,
            scatter src=explicit
        ] table[x=frequency, y=cutUpHeight, meta=diameterToe, col sep=comma] {../../Data/freqCutUpHeightDiam.csv};
        \end{axis}
    \end{tikzpicture}
    \caption{Cut-Up Height vs. Frequency, colored by Diameter Toe.}
\end{figure}


\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            xlabel={Acoustic Intensity},
            ylabel={Modified Ising number},
            grid=major%
        ]
        \addplot table[x=acousticIntensity, y=modifiedIsing, col sep=comma, only marks] {../../Data/accousticIntensityModIsing.csv};
        \end{axis}
    \end{tikzpicture}
    \caption{Acoustic Intensity vs. Modified Ising number.}
\end{figure}



\section{Linear regression for partials distribution}

We use linear regression to study the relationship between the frequencies and
their partials (normalized intensities).
Three heuristic shape functions are used to fit the partials distribution:
\begin{itemize}
    \item \(\mathfrak{p}_\mathsf{lin}(\omega; a, b) = a \omega + b\) -- Linear function.
    \item \(\mathfrak{p}_\mathsf{exp}(\omega; a, b) = a e^{\omega} + b\) -- Exponential function.
    \item \(\mathfrak{p}_\mathsf{log}(\omega; a, b) = a \log(\omega) + b\) -- Logarithmic function.
\end{itemize}

\section{Correction formulas}

Under the heuristic assumption that the Ising number should depend on the flow
rate in a direct proportional way, a first proposed correction formula is
\begin{equation}
    \label{eq:ising_correction}
    \mathsf{I}_\mathsf{corr}(\mathbf{x}) = \mathsf{I}(\mathbf{x})
    (1 + \frac{1}{2} \mathsf{F}(\mathbf{x})),
\end{equation}
where \(\mathsf{F}(\mathbf{x})\) is the flow rate of the wind.

The intent is that, under small flow rates \(|\mathsf{F}(\mathbf{x})| \ll 1\),
the Ising number is approximately equal to the original Ising number.

A second proposed correction formula uses conservation of the flow rate:
\begin{equation}
    \label{eq:ising_correction2}
    \mathsf{I}_\mathsf{corr} = 
    \frac{A_{\mathrm{sys}}}{\omega_0 A_{\mathrm{toe}}}\sqrt{\frac{2 P d}{\rho h^3}}
\end{equation}
where \(A_{\mathrm{sys}}\) is the area of the system --where the wind velocity
has been measured-- and \(A_{\mathrm{toe}}\) is the area of the toe hole.
This integrates small perturbations in value of the Ising number, does not
change its dimensions, and comes from the physical interpretation of the
wind flow as an incompressible fluid.



\chapter[DNN model for Ising number]{A (small) deep neural network model for the Ising number}
\chapterauthor{Gabriel Pinochet-Soto}

Another approach to integrate the factors that fully determine the Ising number
and the partials ratios is to use a deep neural network (DNN).
For this goal, we use the JAX framework~\autocite{2018Jax-1, 2020Optax-1,
2024Flax-1} to implement a ``small'' DNN model.
The architecture of the DNN was made simple due to the limited amount of data
available.

\section{Architecture of the model}

The input features are encoded in a six-dimensional vector
\(\mathbf{x} \in \mathbb{R}^6\), where each component corresponds to the
already mentioned parameters (\textbf{flueDepth}, \textbf{frequency},
\textbf{cutUpHeight}, \textbf{diameterToe}, \textbf{acousticIntensity}).
We perform a \texttt{flax.nnx.BatchNorm} normalization of the input,
inject the output into a larger dimensional space (of hidden layers),
after compressing the input into a nine-dimensional output vector,
which consists of the Ising number and the partials ratios
\(\mathbf{y} \in \mathbb{R}^9\).
We expect \(\mathbf{y}_0 \approx \mathsf{I}(\mathbf{x})\) to be, approximately,
the Ising number, and \(\mathbf{y}_1, \ldots, \mathbf{y}_8\) to be the partials
associated to the \(\mathbf{x}\) observation.
See~\autoref{fig:dnn_architecture} for the architecture of the DNN.

The specific independent softmax-softplus layers are used to ensure outputs
are in the correct range: the Ising number is non-negative, and the partials
ratios are normalized between 0 and 1.
We favor to use hyperbolic tangent activation functions for the hidden layers,
as they are zero-centered and can help with the convergence of the model.
Moreover, it allows for more flexibility in the model, as it can consider
negative values in the hidden layers.

For convenience, we denote the DNN model as \(\mathsf{DNN}: \mathbb{R}^6 \to
\mathbb{R}^9\); in particular, we denote
\(\mathsf{DNN}_\mathsf{I}(\mathbf{x})\in [0, +\infty)\)
as the Ising number prediction associated to the input \(\mathbf{x}\), and
\(\mathsf{DNN}_\mathsf{P}(\mathbf{x}) \in [0, 1]^8\)
as the partials ratios prediction associated to the input \(\mathbf{x}\).

\input{dnn_tikz}

\subsubsection{Some implementation details}

One key aspect of the implementation is the use of activation functions.
There are different types of activation functions, and the criteria for
choosing them is based on the design and goals of the model.

We use hyperbolic tangent activation functions for the hidden layers, as they
are zero-centered, smooth, and can help with the convergence of the model.
\[
    \mathsf{tanh}(x) = \frac{e^x - e^{-x}}{e^x + e^{-x}}.
\]
Concerning the last outputs, we want to recover the Ising number and the
partials ratios.
For the Ising number, we use a softplus activation function, which is defined as
\[
    \mathsf{softplus}(x) = \log(1 + e^x).
\]
This function is a smooth approximation of the ReLU function, and it is always
non-negative, which is suitable for the Ising number prediction.
For the partials ratios, we use a softmax activation function, which is defined
as
\[
    \mathsf{softmax}(\mathbf{y})_j = \frac{e^{y_j}}{\sum_{k=1}^8 e^{y_k}},
\]
where \(\mathbf{y} = (y_1, \ldots, y_8)\) is the output vector of the DNN
associated to the partials ratios.
Notice that we require to add an \emph{extra normalization step} where we divide the
softmax output by the maximum value of the partials ratios, so that the
partials ratios are normalized between 0 and 1, and have the same scale as the
observational data.


\section{Loss function}

The loss function consists of two terms: an squared error term for the
predicted Ising number and the exact Ising number, and a mean squared error term
for the predicted partials ratios and the exact partials ratios.
This is, given observational data \(\{ (\mathbf{x}_i, \mathbf{p}_i)\}_{i=1}^N
\subset \mathbb{R}^6 \times [0,1]^8\), where \(\mathbf{p}_i\) are the
partials ratios (normalized between 0 and 1), and the Ising number formula
\(\mathsf{I}(\mathbf{x})\) is given by~\eqref{eq:ising}, we define the loss
function as follows:
\begin{equation}
    \label{eq:dnn_loss}
    \mathscr{L}_\mathsf{DNN}(\mathbf{W}, \mathbf{b}; \lambda) =
    \frac{1}{N} \sum_{i=1}^N \left(
        \lambda^2 \left( \mathsf{DNN}_\mathsf{I}(\mathbf{x}_i) - \mathsf{I}(\mathbf{x}_i) \right)^2
        +
        \sum_{j=1}^8 \left( \mathsf{DNN}_\mathsf{P}(\mathbf{x}_i)_j - \mathbf{p}_{i,j} \right)^2
    \right),
\end{equation}
where \(\lambda\) is a hyperparameter that controls the relative importance of
the Ising number prediction and the partials ratios prediction.
The hyprerparameter tuning is explained in the next section~\autoref{sec:hyperparameter_tuning}.

\section{Hyperparameter tuning and optimization}
\label{sec:hyperparameter_tuning}

This part was mainly heuristic; we expect to obtain a better model under
more diverse and distributed data.

We use the Optax library~\autocite{2020Optax-1} to optimize the loss function.
Adam~\autocite{2014Kin-1} is used as the optimizer.
A scheduler is used to control the learning rate, which is set to
be a cosine decay function~\autocite{2016LosHut-1}.
We chain the optimizer with a gradient clipping function and a weight decay
regularization term.
The configuration of the hyperparameters is summarized
in~\autoref{tab:dnn_hyperparameters}.

\begin{table}[htbp]
    \centering
    \begin{tabular}{lll}
        \hline
        \textbf{Hyperparameter} & \textbf{Value} & \textbf{Type} \\
        \hline
        \(N_\mathsf{hidden}\) (number of hidden layers) &
        2 & \texttt{int} \\
        \(d_\mathsf{hidden}\) (dimension of hidden layers) &
        16 & \texttt{int} \\
        \(\sigma\) (standard deviation bias initialization) &
        0.325 & \texttt{float} \\
        \(\lambda\) (importance of Ising number prediction) &
        2.5 & \texttt{float} \\
        \(n_\mathsf{decay}\) (number of decay steps) &
        10 & \texttt{int} \\
        \(\alpha_\mathsf{decay}\) (decay convex factor) &
        0.875 & \texttt{float} \\
        \(p_\mathsf{decay}\) (decay power factor) &
        1.125 & \texttt{float} \\
        \(\varepsilon\) (division regularization) &
        1.0e-8 & \texttt{float} \\
        \(\beta_{1,\mathsf{Adam}}\) (Adam first moment) &
        0.95 & \texttt{float} \\
        \(\beta_{2,\mathsf{Adam}}\) (Adam second moment) &
        0.9995 & \texttt{float} \\
        \(\tau_\mathsf{clip}\) (gradient clipping factor) &
        1.0 & \texttt{float} \\
        \(\tau_\mathsf{weight}\) (weight decay factor) &
        1.0e-2 & \texttt{float} \\
        \(\rho\) (weight regularization factor) &
        1.0e-4 & \texttt{float} \\
        \hline
    \end{tabular}
    \caption{Hyperparameters used for the DNN model.}
    \label{tab:dnn_hyperparameters}
\end{table}

\section{Utilization of the model}

We provide a command line interface (CLI) for interacting with the Python
package \texttt{dnnpype}.
The main scripts are \texttt{run\_model}, \texttt{classify\_samples}, and
\texttt{convert\_units}.
\begin{itemize}
    \item \texttt{run\_model} -- Runs the DNN model for training or evaluation.
        It accepts several parameters, such as the number of epochs, batch size,
        learning rate, and data path.
    \item \texttt{classify\_samples} -- Classifies samples to generate a rating
        (and further study) of ``optimal'' partials ratios.
    \item \texttt{convert\_units} -- Small extension of Pint~\autocite{2025Pint-1}
        to convert units of the parameters.
\end{itemize}
See Code~\autoref{code:example} for an example of how to use the DNN model.
On executing Code~\autoref{code:example-run}, 
the model will be trained for 300 epochs with a learning rate of 0.01;
we report a final total average loss of \(0.0393\).
Other statistics are displayed on screen.
A verification step on a test set of the data is performed, and the results are
displayed on screen.

\begin{code}
\captionof{listing}{\texttt{run\_model} CLI Example run}
\label{code:example-run}
\mintinline{bash}{run_model --learning_rate 0.01 --epoch 300 --data_path <path_to_data>},
\end{code}

\include{cli_screen}
\include{cli_samples}
\include{cli_units}

\section{Prospective improvements}

We would like to see more variability in the data, as the current data is 
biased and limited --and justifiedly so, as the client has a limited number of
material to produce the organ pipes.
In order to learn more about the dynamics of the partials (ratios), we
would like to see suboptimal partial ratios.

Other topic of concern is the utilization of the acoustic intensity as a
predictor of the model.
In principle, the acoustic intensity is a measure of the sound level
produced by the pipe --coming from the physics of the sound waves.
This current approach might not be optimal; an attempt to correct and use
the acoustic intensity as a response variable was made, but it impacted the
scaling of the model.
This last statement means to modify the model and the loss function in the
following way: let \(\widetilde{\mathsf{DNN}}: \mathbb{R}^5 \to \mathbb{R}^{10}\)
be the DNN model that takes five input features (the first five parameters 
listed above), and outputs a ten-dimensional vector (the Ising number, the
acoustic intensity, and the eight partials ratios).
Here, $\widetilde{\mathsf{DNN}}_\mathsf{I}(\mathbf{x}) \in [0, +\infty)$ is the Ising
number component, $\widetilde{\mathsf{DNN}}_\mathsf{A}(\mathbf{x}) \in [0, +\infty)$ is
the acoustic intensity component, and
$\widetilde{\mathsf{DNN}}_\mathsf{P}(\mathbf{x}) \in [0, 1]^8$ is the partials ratios
component.
We could modify the loss function~\eqref{eq:dnn_loss} to be
\begin{multline}
    \label{eq:dnn_loss_acoustic}
    \mathscr{L}_{\widetilde{\mathsf{DNN}}}(\mathbf{W}, \mathbf{b}; \lambda) =
    \frac{1}{N} \sum_{i=1}^N \left(
        \lambda^2 \left( \mathsf{DNN}_\mathsf{I}(\mathbf{x}_i) - \mathsf{I}(\mathbf{x}_i) \right)^2
        \right. \\
        \left. + \left( \widetilde{\mathsf{DNN}}_\mathsf{A}(\mathbf{x}_i) - a_i \right)^2
        +
        \sum_{j=1}^8 \left( \mathsf{DNN}_\mathsf{P}(\mathbf{x}_i)_j - \mathbf{p}_{i,j} \right)^2
    \right).
\end{multline}

Along the same lines, it would be interesting to make the current responses
associated to the partials more physical.
Currently (and loosely speaking), out model aims to mimic the behavior of the
intensity and partial distribution measuring tool, while adding the Ising
number constraint.
Note that this artificially decouples the intensity and the partials ratios.
Moreover, the partial distribution, as is notmalized by the maximum partial
intensity, makes the model unable to distinguish the impact of the acoustic
intensity on the partials ratios.
One possible approach is to switch the last layer of the model to only employ
a softmax activation function, and to use the acoustic intensity as a rescaling
factor for the partials ratios: given the acoustic intensity \(A\) and the
partials distribution \(\mathbf{p} = (p_1, \ldots, p_8)\), we could
work and fit the \emph{physical} partials distribution
\(A \cdot \mathbf{p} = (A p_1, \ldots, A p_8)\).

Lastly, we would like to append more data in the current format we are using
(i.e., the \texttt{csv} format).
The current miniapp allows to reuse the models and the data, and we can extend
it to visualize the results (in a graphical way, instead of the CLI TUI).

\chapter{Generalized Additive Models and the Ising number}
\chapterauthor{Meghana Cyanam}

% TODO: Meghana and Daniel had this

\nocite{*} % This will include all references in the bibliography
\printchicagobibliography

% \chapter*{Notes}
% \addcontentsline{toc}{chapter}{Notes}
% \printchicagonotes

% \appendix

\end{document}
